---
title: "Corrected Global Gender Gap Report 2015"
author: "Václav Haisman"
date: "March 1, 2016"
toc: true
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: true
language: english
geometry:
    - a4paper
    - margin=2cm
mainfont: TeX Gyre Schola
...

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(sqldf)
library(tcltk)
library(gsubfn)
library(proto)
library(RSQLite)
library(DBI)

readCsv <- function(fileName, colNames) {
  DF <- read.csv2(fileName, fileEncoding = "UTF-8",
                  sep = ",", quote = "\"", header = FALSE, dec = ".", na.strings = c("—"),
                  col.names = colNames)
  return(DF)
}

linechart <- function(origData, correctedData, chartLabel) {
  frame()
  stripchart(origData, method = "stack",# jitter = 0.1, offset = 1/3,
             vertical = FALSE, add = FALSE,
             at = NULL, xlim = c(0,1), ylim = NULL,
             main = chartLabel, family = "serif", ylab = "", xlab = "ratio",
             log = "", cex.main = 0.8, col="blue", pch=1)
  stripchart(correctedData, method = "stack",# jitter = 0.1, offset = 1/3,
             vertical = FALSE, add = TRUE,
             at = NULL, xlim = c(0,1), ylim = NULL,
             ylab = "", xlab = "",
             log = "", cex.main = 0.8,  col="green", pch=2)
  legend("bottomleft", legend = c("original", "corrected"), col=c("blue", "green"), pch=c(1,2))
}

addMoreInfoColumns <- function(dataFrame) {
  dataFrame$cRank <- rank(-dataFrame$cRatio, na.last = TRUE, ties.method = "min")
  dataFrame$change <- ifelse(dataFrame$cRank <= dataFrame$rank,
                             ifelse(dataFrame$cRank < dataFrame$rank,
                                    sprintf("↑%d", dataFrame$rank - dataFrame$cRank),
                                    "="),
                             sprintf("↓%d", dataFrame$cRank - dataFrame$rank))
  dataFrame <- dataFrame[order(dataFrame$cRank, na.last = TRUE), , drop = FALSE]
  return(dataFrame)
}

subIndexTweak <- function(dataFrame) {
  dataFrame$rank <- rank(-dataFrame$subindex, na.last = TRUE, ties.method = "min")
  dataFrame <- dataFrame[order(dataFrame$rank, na.last = TRUE), , drop = FALSE]
  return(dataFrame)
}

finalIndexTweak <- function(dataFrame) {
  dataFrame$rank <- rank(-dataFrame$index, na.last = TRUE, ties.method = "min")
  dataFrame <- dataFrame[order(dataFrame$rank, na.last = TRUE), , drop = FALSE]
  return(dataFrame)
}

printTable <- function(dataFrame) {
  kable(dataFrame, digits = 2, row.names = FALSE)
}

printSubIndexWeights <- function(dataFrame) {
  kable(dataFrame, digits = 3, row.names = FALSE)
}

```

# Corrected Global Gender Gap Report 2015

TL;DR

:   Go to the [Corrected Global Gender Gap Index](#corrected-gggr-2015) section.

Why am I doing this?

:   Because the [Global Gender Gap Report](http://reports.weforum.org/global-gender-gap-report-2015/) (henceforth GGGR)
    and its indexes does not accurately reflect gender
    gaps that favour females. Ranks and indexes of some of the countries will be significantly different
    if the gap is penalized both when it favours males and when it favours females.

How does this differ from the original report?

:   Whenever there is inequality (gap) affecting males favouring females, the original report
    truncates any ratio at value 1. In this report, table colum cRatio (corrected ratio) is
    calculated as inverse value ($\frac{1}{ratio}$) of the original ratio. There is also new column
    cRank (corrected rank) which shows ranking of the countries using the cRatio column.
    
Why are do some tables have coparison with the original report and some do not?

:   Scraping the original PDF for the data is hard.

How was this PDF produced?

:   This PDF was produced by use of several tools: `pdftohtml`, Perl, R, RStudio and RMarkdown.
    See [Appendix 1](#appendix-1) for details. There is GitHub repository
    [`wilx/fixed-gggr-2015`](https://github.com/wilx/fixed-gggr-2015) that contains all the necessary
    sources and generated files to build this PDF file, except for the original
    GGGR 2015 PDF file.

## Economic participation and opportunity

```{r, echo=FALSE}
labForcePart <- readCsv("Labour-force-participation.csv",
                        c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
labForcePart <- addMoreInfoColumns(labForcePart)

wageEq <- readCsv("Wage-equality-survey.csv",
                  c("country", "survey data", "FM", "FMT", "rank", "cRatio"))
wageEq <- addMoreInfoColumns(wageEq)

estEarnedIncome <- readCsv("Estimated-earned-income.csv",
                   c("country", "female", "male", "female cut-off", "male cut-off", "FMT", "rank"))

legSenMan <- readCsv("Legislators,-senior-officials-and-managers.csv",
                     c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
legSenMan <- addMoreInfoColumns(legSenMan)

techWorkers <- readCsv("Professional-and-technical-workers.csv",
                       c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
techWorkers <- addMoreInfoColumns(techWorkers)

subIndexEcon <- sqldf("SELECT country,
                         (COALESCE(labForcePart.cRatio, 0) * 0.199
                          + COALESCE(wageEq.cRatio, 0) * 0.310
                          + COALESCE(estEarnedIncome.FMT, 0) * 0.221
                          + COALESCE(legSenMan.cRatio, 0) * 0.149
                          + COALESCE(techWorkers.cRatio, 0) * 0.121) AS [subindex]
                      FROM labForcePart
                      JOIN wageEq USING (country)
                      JOIN estEarnedIncome USING (country)
                      JOIN legSenMan USING (country)
                      JOIN techWorkers USING (country)")
subIndexEcon <- subIndexTweak(subIndexEcon)

```

### Labour force participation

````{r, echo=FALSE}
printTable(labForcePart)
````

````{r, echo=FALSE, fig.align='center'}
linechart(labForcePart$FMT, labForcePart$cRatio, "Labour force participation")
````

### Wage equality

````{r, echo=FALSE}
printTable(wageEq)
````

````{r, echo=FALSE, fig.align='center'}
linechart(wageEq$FMT, wageEq$cRatio, "Wage equality")
````


### Estimated earned income

````{r, echo=FALSE}
printTable(estEarnedIncome)
````

### Legislators, senior officials and managers

````{r, echo=FALSE}
printTable(legSenMan)
````

````{r, echo=FALSE, fig.align='center'}
linechart(legSenMan$FMT, legSenMan$cRatio, "Legislators, senior officials and managers")
````


### Professional and technical workers

````{r, echo=FALSE}
printTable(techWorkers)
````

````{r, echo=FALSE, fig.align='center'}
linechart(techWorkers$FMT, techWorkers$cRatio, "Professional and technical workers")
````

### Subindex

With all the corrected ratios calculated above we can now calcualte the Economic Participation and Opportunity Subindex.
According to the original GGGR 2015, the subindex is weighted average of ratios obtained earlier with these weights:

````{r, echo=FALSE}
printSubIndexWeights(
  data.frame(ratio = c("Ratio: female labour force participation over male value",
                       "Wage equality between women and men for similar work (converted to female-over-male ratio)",
                       "Ratio: female estimated earned income over male value",
                       "Ratio: female legislators, senior officials and managers over male value",
                       "Ratio: female professional and technical workers over male value"),
             weight = c(0.199, 0.310, 0.221, 0.149, 0.121)))
````

The resulting corrected country ranks and subindex values are as follows:

````{r, echo=FALSE}
printTable(subIndexEcon)
````


## Education attainment

```{r, echo=FALSE}
primEd <- readCsv("Enrolment-in-primary-education.csv",
                  c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
primEd <- addMoreInfoColumns(primEd)

secEd <- readCsv("Enrolment-in-secondary-education.csv",
                 c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
secEd <- addMoreInfoColumns(secEd)

terEd <- readCsv("Enrolment-in-tertiary-education.csv",
                 c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
terEd <- addMoreInfoColumns(terEd)

literacy <- readCsv("Literacy-rate.csv",
                    c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
literacy <- addMoreInfoColumns(literacy)

subIndexEdu <- sqldf("SELECT country,
                         (COALESCE(primEd.cRatio, secEd.cRatio, terEd.cRatio, 0) * 0.191
                          + COALESCE(secEd.cRatio, terEd.cRatio, 0) * 0.459
                          + COALESCE(terEd.cRatio, 0) * 0.230
                          + COALESCE(literacy.cRatio, 0) * 0.121) AS [subindex]
                      FROM primEd
                      JOIN secEd USING (country)
                      JOIN terEd USING (country)
                      JOIN literacy USING (country)")
subIndexEdu <- subIndexTweak(subIndexEdu)

```

### Literacy

````{r, echo=FALSE}
printTable(literacy)
````

````{r, echo=FALSE, fig.align='center'}
linechart(literacy$FMT, literacy$cRatio, "Literacy")
````


### Primary education enrollment

````{r, echo=FALSE}
printTable(primEd)
````

````{r, echo=FALSE, fig.align='center'}
linechart(primEd$FMT, primEd$cRatio, "Primary education enrollment")
````


### Secondary education enrollment

````{r, echo=FALSE}
printTable(secEd)
````

````{r, echo=FALSE, fig.align='center'}
linechart(secEd$FMT, secEd$cRatio, "Secondary education enrollment")
````

### Tertiary education enrollment

````{r, echo=FALSE}
printTable(terEd)
````

````{r, echo=FALSE, fig.align='center'}
linechart(terEd$FMT, terEd$cRatio, "Tertiary education enrollment")
````

### Subindex

With all the corrected ratios calculated above we can now calcualte the Educational Attainment Subindex.
According to the original GGGR 2015, the subindex has the following weights:

````{r, echo=FALSE}
printSubIndexWeights(
  data.frame(ratio = c("Ratio: female literacy rate over male value",
                       "Ratio: female net primary enrolment rate over male value",
                       "Ratio: female net secondary enrolment rate over male value",
                       "Ratio: female gross tertiary enrolement ratio over male value"),
             weight = c(0.191, 0.459, 0.230, 0.121)))
````

The resulting corrected country ranks and subindex values are as follows:

````{r, echo=FALSE}
printTable(subIndexEdu)
````



## Health and survival

```{r, echo=FALSE}
birthRatio <- readCsv("Sex-ratio-at-birth.csv",
                      c("country", "female", "male", "FMT", "rank"))

healthyLife <- readCsv("Healthy-life-expectancy.csv",
                       c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
healthyLife <- addMoreInfoColumns(healthyLife)

subIndexHealth <- sqldf("SELECT country,
                         (COALESCE(birthRatio.FMT, 0) * 0.693
                          + COALESCE(healthyLife.cRatio, 0) * 0.307) AS [subindex]
                      FROM birthRatio
                      JOIN healthyLife USING (country)")
subIndexHealth <- subIndexTweak(subIndexHealth)
```

### Sex ratio at birth

````{r, echo=FALSE}
printTable(birthRatio)
````

````{r, echo=FALSE, fig.align='center'}
linechart(birthRatio$FMT, birthRatio$FMT, "Sex ratio at birth")
````

### Healthy life expectancy

````{r, echo=FALSE}
printTable(healthyLife)
````

````{r, echo=FALSE, fig.align='center'}
linechart(healthyLife$FMT, healthyLife$cRatio, "Sex ratio at birth")
````

### Subindex

With all the corrected ratios calculated above we can now calcualte the Health and Survival Subindex.
According to the original GGGR 2015, the subindex has the following weights:

````{r, echo=FALSE}
printSubIndexWeights(
  data.frame(ratio = c("Sex ratio at birth (converted to female-over-male ratio)",
                       "Ratio: female healthy life expectancy over male value"),
             weight = c(0.693, 0.307)))
````

The resulting corrected country ranks and subindex values are as follows:

````{r, echo=FALSE}
printTable(subIndexHealth)
````



## Political empowerment

```{r, echo=FALSE}
womenInParliament <- readCsv("Women-in-parliament.csv",
                             c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
womenInParliament <- addMoreInfoColumns(womenInParliament)

womenMinisters <- readCsv("Women-in-ministerial-positions.csv",
                          c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
womenMinisters <- addMoreInfoColumns(womenMinisters)

femaleHeadOfState <- readCsv("Years-with-female-head-of-state.csv",
                             c("country", "female", "male", "FM", "FMT", "rank", "cRatio"))
femaleHeadOfState <- addMoreInfoColumns(femaleHeadOfState)

subIndexPol <- sqldf("SELECT country,
                         (COALESCE(womenInParliament.cRatio, 0) * 0.310
                          + COALESCE(womenMinisters.cRatio, 0) * 0.247
                          + COALESCE(femaleHeadOfState.cRatio, 0) * 0.443) AS [subindex]
                      FROM womenInParliament
                      JOIN womenMinisters USING (country)
                      JOIN femaleHeadOfState USING (country)")
subIndexPol <- subIndexTweak(subIndexPol)
```


### Women in parliament

````{r, echo=FALSE}
printTable(womenInParliament)
````

````{r, echo=FALSE, fig.align='center'}
linechart(womenInParliament$FMT, womenInParliament$cRatio, "Women in parliament")
````

### Women in ministerial positions

````{r, echo=FALSE}
printTable(womenMinisters)
````

````{r, echo=FALSE, fig.align='center'}
linechart(womenMinisters$FMT, womenMinisters$cRatio, "Women in ministerial positions")
````

### Years with female head of state

````{r, echo=FALSE}
printTable(femaleHeadOfState)
````

````{r, echo=FALSE, fig.align='center'}
linechart(femaleHeadOfState$FMT, femaleHeadOfState$cRatio, "Years with female head of state")
````

### Subindex

With all the corrected ratios calculated above we can now calcualte the Political Empowerment Subindex.
According to the original GGGR 2015, the subindex has the following weights:

````{r, echo=FALSE}
printSubIndexWeights(
  data.frame(ratio = c("Ratio: females with seats in parliament over male value",
                       "Ratio: females at ministerial level over male value",
                       "Ratio: number of years with a female head of state (last 50 years) over male value"),
             weight = c(0.310, 0.247, 0.443)))
````

The resulting corrected country ranks and subindex values are as follows:

````{r, echo=FALSE}
printTable(subIndexPol)
````


## Corrected Global Gender Gap Index {#corrected-gggr-2015}

````{r, echo=FALSE}
finalIndex <- sqldf("SELECT country,
                         ((COALESCE(subIndexEcon.subindex, 0)
                           + COALESCE(subIndexEdu.subindex, 0)
                           + COALESCE(subIndexHealth.subindex, 0)
                           + COALESCE(subIndexPol.subindex, 0))
                          / 4.0) AS [index]
                      FROM subIndexEcon
                      JOIN subIndexEdu USING (country)
                      JOIN subIndexHealth USING (country)
                      JOIN subIndexPol USING (country)")
finalIndex <- finalIndexTweak(finalIndex)
````

The final index is unweighted average of the four subindexes calculated above:

````{r, echo=FALSE}
printTable(finalIndex)
````

````{r, echo=FALSE}
stripchart(finalIndex$index, method = "stack",# jitter = 0.1, offset = 1/3,
           vertical = FALSE, add = FALSE,
           at = NULL, xlim = c(0,1), ylim = NULL,
           main = "Corrected Global Gender Gap Index", family = "serif", ylab = "", xlab = "ratio",
           log = "", cex.main = 0.8, col="blue", pch=1)
````

To compare these corrected results with the original GGGR 2015, you will have to consult the original report PDF.

# Appendix 1 {#appendix-1}

The following Perl script was used to scrape GGGR PDF file and produce separate CSV for each table:

```perl
use strict;
use XML::XPath;
use XML::XPath::XMLParser;
use Text::CSV::Encoded;
use IO::File;
use IO::Handle;
use Regexp::Common;
use Data::Dumper;

binmode(STDOUT, ":encoding(UTF-8)");

my $inputf = IO::File->new();
$inputf->open("gggr.xml", "<");
$inputf->binmode(":encoding(UTF-8)");
my $xp = XML::XPath->new($inputf);
my $pages = $xp->find('//page');

foreach my $page ($pages->get_nodelist) {
    my @table_title_nodes = $page->find('./text[@font = \'1\']')->get_nodelist;
    my $table_title = $table_title_nodes[0]->string_value;
    print STDERR "table title: ", $table_title, "\n";
    my @value_nodes = $page->find('./text[@font = \'5\' and @height = \'10\']'
                                  . '| ./text[@font = \'7\' and @height = \'11\']')->get_nodelist;
    my @value_text = map {$_->string_value} @value_nodes;

    # Count from 2nd element all elements that are numbers. This will be the
    # column count for this table.

    my $columns;
    for ($columns = 0; $value_text[$columns + 1] =~ /^\d/; ++$columns)
    {}
    ++$columns;
    print STDERR "detected ", $columns, " columns\n";

    # Create separate CSV file for each table.

    my $file_name = $table_title;
    $file_name =~ s/.*?:\s*(\S.*)/$1/;
    $file_name =~ s/\s+/-/g;
    my $table_name = $file_name;
    $file_name .= '.csv';
    my $csv = Text::CSV::Encoded->new({encoding => undef, binary => 1});
    my $f = IO::File->new();
    $f->open("$file_name", ">");
    #$f->binmode(":encoding(UTF-8)");

    # Output all rows.

    # Each record specifies columns of F/M ratio, F/M ratio truncated and
    # rank.

    my %table_info = (
        'Enrolment-in-primary-education' => [3, 4, 5],
        'Enrolment-in-secondary-education' => [3, 4, 5],
        'Enrolment-in-tertiary-education' => [3, 4, 5],
        'Estimated-earned-income' => undef,
        'Healthy-life-expectancy' => [3, 4, 5],
        'Labour-force-participation' => [3, 4, 5],
        'Legislators,-senior-officials-and-managers' => [3, 4, 5],
        'Literacy-rate' => [3, 4, 5],
        'Professional-and-technical-workers' => [3, 4, 5],
        'Sex-ratio-at-birth' => undef,
        'Wage-equality-survey' => [2, 3, 4],
        'Women-in-ministerial-positions' => [3, 4, 5],
        'Women-in-parliament' => [3, 4, 5],
        'Years-with-female-head-of-state' => [3, 4, 5]
        );

    my $size = scalar @value_text;
    for (my $i = 0; $i < $size; $i += $columns) {
        my @row = @value_text[$i..($i + $columns - 1)];
        #print STDERR "row: ", Dumper(\@row), "\n";

        if ($table_name eq 'Estimated-earned-income') {
            $row[1] =~ s/,//g;
            $row[2] =~ s/,//g;
            $row[3] =~ s/,//g;
            $row[4] =~ s/,//g;
        }

        # Try to compute ratio corrected for instances where F/M ratio > 1.
        if (exists $table_info{$table_name}
            && defined $table_info{$table_name}) {
            my @cols = @{$table_info{$table_name}};
            my $fmr = $row[$cols[0]];
            my $orig_fmr = $fmr;
            my $diff = 0;
            if ($fmr =~ /$RE{num}{real}/) {
                $fmr += 0.0;
                if ($fmr > 1) {
                    $fmr = 1.0 / $fmr;
                }
                $diff = $fmr - $orig_fmr;
                if ($diff != 0) {
                    print STDERR ("", $row[0], ": ", $orig_fmr, " -> ", (sprintf "%5.2f", $fmr),
                                  " (delta ", (sprintf "%5.2f", $diff), ")\n");
                }
            }
            push @row, $fmr;
            #push @row, $diff;
        }

        $csv->print($f, \@row);
        print $f "\n";
    }
}
```

The script uses `gggr.xml` file as its input. This file was produced by the following command:

```bash
pdftohtml -f $((52+8)) -l $((65+8)) -c -s -noframes -xml -enc UTF-8  gggr2015.pdf gggr
```

The `gggr2015.pdf` file can be [downloaded](http://reports.weforum.org/global-gender-gap-report-2015/) from World Economic Forum web site.
